A Dearth of Young Stars and Planets in the Kepler Field

### 2024-05-27 paper/ms.tex
update ms

The stellar ages 
are given in Table~\ref{tab:planets} for the known planet-hosts,
and in Table~\ref{tab:stars} for all Kepler stars.  Figure~\ref{fig:hist_tgyro} shows the gyrochone
age distributions of all the stars (left column) and the known
planet-hosts (middle column).  These ``histograms'' are sampled from
the posterior probability distributions by drawing ten random samples
from each posterior, and then computing the re-weighted
histogram of the resulting samples.  The uncertainties assume Poisson
statistics.  We truncated the plots near $4$\,Gyr since this is the
cluster age out to which our interpolation-based method is calibrated.
The overall slope yields a dearth of young stars relative to the expectation
of a uniform age distribution.
%In an alternative coinage, this tendency for fewer young
%stars to be present could also be dubbed a ``demographic cliff'',
%in reference to the phenomenon from human demography.

The age distributions of Kepler stars and planet hosts are quite
similar.  A simple diagnostic of whether the two
distributions are drawn from the same underlying distribution is the
Kolmogorov-Smirnov test;  we calculated this statistic in a manner
that accounts for the Poisson uncertainties by performing 1000 random
draws of 70\% of the stars from each sample.  The resulting 
range of $\log_{\rm 10}p$-values (2.5$^{\rm th}$ to 97.5$^{\rm th}$ percentile)
spanned -3.9 to -1.2 for the K-dwarfs only,
and -5.2 to -2.2 for all Kepler stars.  This degree of similarity suggests that
the ``common wisdom'' that young transiting planets are hard to detect
because their host stars are photometrically variable requires some
nuance.  If true, this statement only holds for the very youngest
stars, since at the age resolution of this study, there appears to be
little to no selection bias against sub-gigayear planets.

>>>>

The stellar ages 
are given in Table~\ref{tab:planets} for the known planet-hosts,
and in Table~\ref{tab:stars} for all Kepler stars.  Figure~\ref{fig:hist_tgyro} shows the rotation-based
age distributions of all the stars (left column) and the known
planet-hosts (middle column).  These ``histograms'' are sampled from
the posterior probability distributions by drawing ten random samples
from each posterior, and then computing the re-weighted
histogram of the resulting samples.  The uncertainties assume Poisson
statistics.  We truncated the plots near $4$\,Gyr since this is the
cluster age out to which our interpolation-based method is calibrated.
The overall slope yields a dearth of young stars relative to the expectation
of a uniform age distribution.
%In an alternative coinage, this tendency for fewer young
%stars to be present could also be dubbed a ``demographic cliff'',
%in reference to the phenomenon from human demography.

The age distributions of Kepler stars and KOI hosts in Figure~\ref{fig:hist_tgyro} appear
similar.  One diagnostic for whether the two
distributions are drawn from the same underlying distribution is the
Kolmogorov-Smirnov test;  we calculated this statistic in a manner
that accounts for the Poisson uncertainties by performing 1000 random
draws of 70\% of the stars from each sample.  The resulting 
range of $\log_{\rm 10}p$ (2.5$^{\rm th}$ to 97.5$^{\rm th}$ percentile)
spanned -3.9 to -1.2 for the K-dwarfs only,
and -5.2 to -2.2 for all Kepler stars.  This degree of similarity suggests 
that although weak differences may be present,
the ``common wisdom'' that young transiting planets are hard to detect
because their host stars are photometrically variable requires some
nuance.  If true, this statement only holds for the very youngest
stars, where we see only a marginal underabundance in the KOI host age
distribution relative to that of the overall stellar sample.
At the age resolution of this study, 
the selection bias against sub-gigayear planets seems small.
